TBD

### 2020-11-09 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Edits and elaboration
derivation summary, derivation exposition edits, fundamentality, reorganization of sections

\usepackage{amsmath, hyperref, enumerate}
\usepackage{amsfonts}
\usepackage{graphicx}

@@

\item motivate everything upfront: going to show that Hamiltonian mechanics corresponds to two conditions being satisfied, while Lagrangian mechanics requires a three condition
\item define the two assumptions, and give motivation
\item figuring out which mathematical structures are required for systems that satisfy certain physical conditions
\item go over the upshot of the philosophical argument, specifically how this avoids the worries about measuring the amount of structure that plagued North (i.e. Swanson and Halvorson paper)
\item Then, give the derivation, as gentle and well-motivated as possible. But would be ideal if readers could skip or skim this section and still get the philosophical upshot, relatively early on.

@@

TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.

Philosophers have recently shown that we can interpret Lagrangian and Hamiltonian mechanics as theoretically equivalent in a precise mathematical sense. Within a subclass of classical mechanical systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh2017}{Barrett2018}. Categorical equivalence provides a natural interpretive criterion for theoretical equivalence because it yields mutual inter-translatability between theory formulations. At first glance, this equivalence seems to motivate the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain. This view would undermine prior debates that have privileged either Hamiltonian mechanics or Lagrangian mechanics on physical or metaphysical grounds \parencites[]{North2009}{Curiel2014}. Here, we will argue that there are indeed physical grounds for privileging Hamiltonian mechanics over Lagrangian mechanics, their categorical equivalence notwithstanding. Compared to Hamiltonian systems, an additional physical condition must be satisfied for a classical system to be Lagrangian. Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this conceptual difference has interpretive consequences even within the hyper-regular domain. Furthermore, the construction at the heart of this argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density, along with a novel interpretation of the principle of stationary action within the Hamiltonian formalism.

@@

\subsection{Define Classical Mechanics}

Talk about dissipative systems.

Throughout, we will focus only on non-dissipative classical systems, i.e. conservative systems. For in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $. This requires that the system is conservative [why exactly?], which is equivalent to the system being deterministic and reversible. Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
% Dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential

\subsection{Derivation}

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{Diagram.png}
\end{figure}

Overall summary: A classical system is an infinitesimally reducible system. The state space of these systems can be modeled by a symplectic manifold (the physical assumption of infinitesimal reducibility corresponds to the mathematical structure of symplectic manifolds). Classical Hamiltonian systems satisfy the additional conditions of having deterministic and reversible evolutions. These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it a contact manifold. Classical Lagrangian systems satisfy a further assumption: keeping track of the trajectories in physical space is equivalent to tracking the evolution of states in phase space. Note that it is not always true that we can determine the evolution in phase space from the trajectory in physical space. For instance, knowledge of the trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from its trajectory alone. Hence, the set of classical systems that satisfy this further physical condition is a proper subset of the classical systems that satisfy the assumptions for Hamiltonian mechanics.

\subsubsection{Infinitesimal reducibility}

@@

\begin{quotation}
Infinitesimal reducibility assumption: the state of the system is reducible to the state of its infinitesimal parts. That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the states of those parts, and so on.
\end{quotation}

@@

Let $\mathcal{C}$ be the state space for the whole system. Any point $c \in \mathcal{C}$ represents a particular state of the whole system. We define a classical \textit{particle} as the limit of recursive subdivision. A classical system then consists of a collection of classical particles. Let $\mathcal{S}$ be the state space for the particles [all of the particles constituting the whole system? or for each particle individually?]. We assume that this state space $\mathcal{S}$ is a manifold. Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts. That is, the state of the whole system is a distribution over its infinitesimal parts. If $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space $\mathcal{S}$. 
%[is region U in the state space C or S. I assume in S.].

On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function, i.e. it must depend on the point [the point $c $? ] and not on the coordinates. This is because the distribution is a function only of the states, rather than of the particular choice of state variables, i.e. coordinates. [Explain why these two physical conditions hold for the distribution. Also, explain what it means to transform as a density.] We will now show that classical phase space (i.e. a symplectic manifold) is the only space that satisfies these constraints on the distribution.

We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state (in manifold terminology, these are known as coordinates). We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit. For instance, $q$ might be position, defining units of meters. Starting with the simplest case, suppose that one coordinate suffices to define the unit system for the particle state space. Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number. We will show that, in fact, there can only be a single $k^i$ coordinate in this case, conjugate to the single $q$ coordinate. Our derivation assumes---on physical grounds---that we can (i) arbitrarily change the unit defined by $q$ without changing the density $\rho_c$ and that (ii) $q$ fully defines the units for all state variables. 
%[Give a concrete example of a state variable defining a unit, along with physical motivation for why state variables play the role of defining units (since we are treating this as a physically grounded assumption).]
% How many $k^i$ can we have?

[We need to introduce the definition of state variable evolution $s$, since we appeal to it below. I take it that the definition we are working with is that future states are labeled by the original state variable values (i.e. transported state variables)? Section IV.C of long paper.]

Now, suppose that we change the units by transforming the coordinate $q$ to a new coordinate $\hat{q}$ that is a function of the old coordinate: $\hat{q}=\hat{q}(q)$. Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$. We have $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$. Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must be equal to 1. Suppose there is only one state variable, so that there are no $k^i$ variables. Then we would have $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right| = 1$. But this would mean the unit change cannot be arbitrary. Therefore we must have two or more variables, so that there is at least one $k^i$ variable. The requirement that the Jacobian determinant equals one places a constraint on how the state variables $k^i$ must change to compensate the change from $q$ to $\hat q$. Note that, since $\hat{q}$ depends only on $q$, the Jacobian is an upper triangular block matrix, so $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$ [it might be worthwhile to spell this out in more detail].  For the Jacobian determinant to equal one, we must therefore have $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$. This places a constraint only on the determinant of the transformation. Suppose there are three or more variables. This constraint would then be insufficient to recover the transformation uniquely, and therefore $\hat q$  would not fully define the units for all other state variables (contrary to assumption). This means there must be exactly two state variables: $q$ and a single $k$. So in this case, coordinate-independent areas and densities can only be defined on a two-dimensional manifold.

@@

We can now show that $\mathcal{S}$ is a symplectic manifold. Invariant densities can only be assigned to infinitesimal areas. Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$ [clarify which phase space $\Sigma$ belongs to]. Because degrees of freedom are---at least locally---independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom. This means the volume form is given by $\omega^n$. This cannot be degenerate, or some volumes would not contain states [motivate this claim or]. Therefore $\omega$ itself cannot be degenerate. As the degrees of freedom are independent, the number of states on a surface does not change if we translate it across independent degrees of freedom. If we imagine a parallelepiped, the integral over the surface must be zero (the integrals over opposite sides are equal and opposite). Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold. By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.

The conclusion is that, under the infinitesimal reducibility assumption, the state space of the system is a coordinate-invariant distribution over the state space of the infinitesimal parts [what does it mean for the state space $\mathcal{C} $ to itself be a distribution ?]. If the state space of the particles is a manifold, then it is a symplectic manifold, which is the only type of manifold that supports coordinate-invariant distributions [do we have a reference to substantiate the ``only" claim? Has the proof above really shown that this is the only kind of manifold that would support a coordinate-invariant distribution].

\subsubsection{Deterministic and reversible evolution}

@@

\begin{quotation}

Deterministic and reversible evolution assumption: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.

@@

[Should we footnote that these two assumptions leave out some classical systems, e.g. Norton's dome, Earman's space invaders?]

In our case, this applies to both the state of the whole system and to the state of its parts. Let $\lambda : \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle. Under the assumption, this will be uniquely identified by the initial state $s_0 = \lambda(t_0)$ at the initial time $t_0$. [is $s$ an evolution that transports the state variables, i.e. labeling future states by the original state variables so that the initial time is transported to itself? I think we need to explain more about how evolutions are defined]. Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial particle [what exactly does `initial particle' mean? Do we mean the initial state of the composite system or just a particular particle?] at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e. the density remains the same throughout the evolution. In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa [could two different particles occupy the same initial state?].

>>>>

\usepackage{amsmath, hyperref, enumerate, comment}
\usepackage{amsfonts}
\usepackage{graphicx}

@@

\item motivate everything upfront: going to show that Hamiltonian mechanics corresponds to two conditions being satisfied, while Lagrangian mechanics requires a third condition
\item define the two assumptions, and give motivation
\item figuring out which mathematical structures are required for systems that satisfy certain physical conditions
\item go over the upshot of the philosophical argument, specifically how this avoids the worries about measuring the amount of structure that plagued North (i.e. Swanson and Halvorson paper)
\item Then, give the derivation, as gentle and well-motivated as possible. But would be ideal if readers could skip or skim this section and still get the philosophical upshot, relatively early on.

@@

% TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.

Philosophers have recently shown that we can interpret Lagrangian and Hamiltonian mechanics as theoretically equivalent in a precise mathematical sense. Within a subclass of classical mechanical systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh}{Barrett2}. Categorical equivalence provides a natural interpretive criterion for theoretical equivalence because it yields mutual inter-translatability between theory formulations. At first glance, this equivalence seems to motivate the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain. This equivalence undermines prior positions that have privileged either Hamiltonian mechanics or Lagrangian mechanics on physical or metaphysical grounds \parencites[]{North}{Curiel}. Here, we will argue that there are indeed physical grounds for privileging Hamiltonian mechanics over Lagrangian mechanics, their categorical equivalence notwithstanding. Compared to Hamiltonian systems, an additional physical condition must be satisfied for a classical system to be Lagrangian. Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this difference has interpretive consequences even within the hyper-regular domain. Furthermore, the construction at the heart of this argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density, along with a novel interpretation of the principle of stationary action within the Hamiltonian formalism.

@@

\section{Defining `classical mechanical systems'}

Provide any relevant background on classical mechanics, specifically to clarify/define terminology. Might be able to just put this in the introduction proper, or the start of the next section. Make explicit how our terminology relates to other standard terminology in the philosophical literature.

A classical Hamiltonian system is any system whose phase space is a symplectic manifold and that evolves according to Hamiltonian evolution (i.e. according to a symplectomorphism).

Throughout, we will focus only on non-dissipative classical systems, i.e. conservative systems. For in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $. This requires that the system is conservative, which is equivalent to the system being deterministic and reversible. Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
% Dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential

\section{Privileging Hamiltonian mechanics}

Our framework develops a correspondence between three physical conditions and their associated mathematical formalisms. In this way, we straightforwardly interpret the physical significance of these mathematical formalisms. First, a system is \textit{infinitesimally reducible} provided that the states of its constituent, infinitesimal particles completely specify the overall state of the system. When this condition is satisfied, we can define a distribution $\rho$ over particle states. Furthermore, since particle states are independent of a choice of units, this distribution must be invariant. We show in Section~\ref{infinitesimal} that any infinitesimally reducible system has a phase space given by a symplectic manifold. The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.

Secondly, a system has \textit{deterministic and reversible evolution} provided that the present state of the system uniquely determines both its future and past states. This condition entails that the preceding invariant distribution is in fact a conserved phase space density. We can thus model such systems using Hamiltonian mechanics, defined on a phase space with a symplectomorphism. These two conditions thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution. Section~\ref{deterministic} provides this derivation in detail.

Finally, a system satisfies a further condition---motion equivalence---when its trajectory in physical space determines and is determined by its evolution in phase space. This condition leads to a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics in the hyper-regular domain (Section~\ref{motion}). Hence, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics. To see this, note that some systems fail to satisfy the condition of motion equivalence, i.e. it is not always true that we can determine the evolution of a system in phase space from its trajectory in physical space. For instance, knowledge of the trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from its trajectory alone. Hence, the set of classical systems that satisfy this further physical condition is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics. On these grounds, we will argue in Section~\ref{fundamentality} that Hamiltonian mechanics is more fundamental than Lagrangian mechanics.

% Overall summary: A classical system is an infinitesimally reducible system. The state space of these systems can be modeled by a symplectic manifold (the physical assumption of infinitesimal reducibility corresponds to the mathematical structure of symplectic manifolds). Classical Hamiltonian systems satisfy the additional conditions of having deterministic and reversible evolutions. These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it a contact manifold. Classical Lagrangian systems satisfy a further assumption: keeping track of the trajectories in physical space is equivalent to tracking the evolution of states in phase space.

Figure~\ref{diagram} summarizes our characterization of Hamiltonian and Lagrangian mechanics, in terms of physical conditions and corresponding mathematical structures. The left column lists the three physical conditions: infinitesimal reducibility, determinism/reversibility, and motion equivalence. Each condition entails physically significant properties, described in the middle column. Each of these physical properties corresponds to a formal mathematical framework, described in the right column. Section~\ref{derivation} justifies the logical connection between these physical conditions and corresponding mathematics. Before delving into this derivation, Section~\ref{fundamentality} discusses its philosophical significance.

\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{Diagram.png}
\caption{Physical conditions and corresponding properties and formalisms}
\label{diagram}
\end{figure}

\subsection{Fundamentality and structure}
\label{fundamentality}

In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by ``fundamental." Of the various notions of fundamentality, there is at least the following non-mysterious definition: a formalism is more fundamental provided that it is logically more general. In this case, the Lagrangian framework is less general because it requires that an additional condition hold, compared to the Hamiltonian framework. Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework. In particular, the assumption of motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general. This notion of fundamentality leads as well to a concomitant notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied. In Section~\ref{distribution}, we will discuss another sense in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.
% Way to motivate that the motion equivalence assumption further constrains the Hamiltonian: the Hamiltonian already defines a relationship between conjugate momentum and velocity (through ($dq/dt = \partial H / \partial p$) ). Motion equivalence provides a further constraint on this relationship, and hence a further constraint on the Hamiltonian. \\

% As defined, our notions of fundamentality and structure are not metaphysical in nature.

[Explain how our notion of comparative structure avoids the objections of Halvorson/Swanson against North's comparative notion of amount of structure.] \\

We do not endorse a metaphysical notion of fundamentality, on which a formalism is more fundamental provided that it is more joint-carving. For instance, North provides the following definition of structure: ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" \parencites*[66]{North}. While Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{distribution}), we will also argue that classical mechanics deploys physically significant representations that are not intrinsic to mechanical systems. Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question. Whether Hamiltonian phase space structure is metaphysically fundamental does not matter for the questions we seek to answer. Regardless of its metaphysical status, Hamiltonian mechanics is epistemically, conceptually, and physically more fundamental.

% Part of the following paragraph seems redundant given the preceding. Think about how to combine.
Methodologically, we agree with North that it is important to identify invariants and more generally ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}. Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North}. However, whereas North proceeds mainly from the mathematical structure to claims about physical significance, we seek to interpret the mathematics by beginning with physically significant assumptions. We seek to determine which physical assumptions are necessary and sufficient for supporting a corresponding mathematical structure. This does not entail that additional mathematical structure is necessarily physically insignificant or surplus structure, just that it may correspond to an additional physically significant assumption that might apply to the system under consideration. As we will see, this characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system. Hence, whereas North views the Lagrangian structure as surplus or redundant (``excess, superfluous structure"), we take the Lagrangian structure as unnecessary but still physically significant, corresponding to an additional physical assumption that applies to some classical systems.

Furthermore, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references]. Each of these frameworks has different pragmatic costs and benefits. Hence, unlike North, we are comfortable acknowledging that the mathematical structure that you ``need" might not be unique. There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions. Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others. And in some sense, each of these formalisms references some ``unphysical" coordinates or representational means. Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics. In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.
% One con of this extended phase space formulation: the Hamiltonian constraint is both a constraint and the generator of time evolution; hence, it must be defined in the neighborhood of the actual states, so the values it provides outside of these neighborhoods should be unphysical, yet they encode the equations of motion.
% Should we remark on the claim that the only structure we need is whatever is sufficient to define densities locally? Hence, we need just the symplectic form but not the cotangent bundle?
% E.g. in the extended phase space formulation, the equations of motion come from derivatives of the Hamiltonian constraint over the variables, but this constraint is defined on unphysical points. These unphysical points are plausibly a mathematical artifact, although they supply a nice tool/method for understanding physically significant aspects of the representation.

% North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory. Structure comprises those quantities that remain intact when we transform one allowable coordinatization of the laws into another" [78]

% North definition of structure: ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" [66]. So presumably, North would view the invariant phase space density as part of this fundamental structure.

% North's epistemic method for figuring out the structure: ``look at what is invariant under allowable transformations"; this tells us what the structure is [73].

\section{Derivation}
\label{derivation}

Sections~\ref{infinitesimal}-\ref{deterministic} show how to derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution. Section~\ref{motion} shows how systems that satisfy the further condition of motion equivalence can be treated within the formalism of Lagrangian mechanics.

\subsection{Infinitesimal reducibility}
\label{infinitesimal}

@@

\begin{quotation}
\noindent
\textit{Infinitesimal reducibility}: the state of the system is reducible to the state of its infinitesimal parts. That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the states of those parts, and so on.
\end{quotation}

@@

Let $\mathcal{C}$ be the state space for the whole system. Any point $c \in \mathcal{C}$ represents a particular state of the whole system. We define a classical \textit{particle} as the limit of recursive subdivision. A classical system then consists of a collection of classical particles. Let $\mathcal{S}$ be the state space for the particles, so that each point in $\mathcal{S}$ describes a possible state of a classical particle. We assume that this state space $\mathcal{S}$ is a manifold. Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts. That is, the state of the whole system is a distribution over its infinitesimal parts. This distribution describes the proportion of classical particles in any particular state. Formally, if $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space $\mathcal{S}$. 
%[is region U in the state space C or S. I assume in S.].

We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state (in manifold terminology, these are known as coordinates). Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$, yielding a real-valued vector for any given particle state, with $n$-many degrees of freedom. We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit. For instance, $q$ might be position, defining units of meters. On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function. This is because the distribution is a function only of the states, rather than of the particular choice of state variables, i.e. coordinates. To transform as a density means that $\rho_c$ changes as a function of the Jacobian determinant that relates the new coordinates to the old coordinates. To transform as a scalar means that the distribution $\rho_c$ is invariant under all coordinate transformations, so that the Jacobian determinant must in fact equal one. Hence, $\rho_c$ is a function only of the point $c$ and does not depend on the choice of coordinates. We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold. In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a density and scalar function. % Classical phase space (i.e. a symplectic manifold) is the only space that satisfies these constraints on the distribution.

The argument begins with the simplest case: suppose that a single coordinate $q $ suffices to define the unit system for the particle state space. Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number. We will show that, in fact, there can only be a single $k^i$ coordinate in this case, conjugate to the single $q$ coordinate. Assuming that the particle density $\rho_c$ is a scalar, it follows that we can arbitrarily change the unit defined by $q$ without changing the density $\rho_c$. 
% We further assume that $q$ fully defines the units for all state variables. [Is there a way to physically motivate this assumption, beyond this being a constitutive role of the position-like state variables?]

[We might be able to shorten part of the following argument simply by noting that the Jacobian determinant must equal unity, as explained verbally in a prior paragraph. The preceding verbal justification might be sufficient, i.e. that the distribution transforms as a density and a scalar. We could then focus on the trickier part of showing that there is only a single $k^i$ variable, based on the form of the Jacobian determinant.]

Now, suppose that we change the units by transforming the coordinate $q$ to a new coordinate $\hat{q}$. Since---by assumption---$q$ fully defines the unit system, $\hat{q}$ must do so as well, making it a function just of the old coordinate: $\hat{q}=\hat{q}(q)$. Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$. As defined, $\rho_c$ is a function from the particle state space $\mathcal{S}$ to the reals. We can define $\rho_c$ on the units by taking advantage of an invertible function $s: \mathbb{R}^n \to \mathcal{S}$ mapping a particular value of the units $\hat{\xi}^b$ to a point in the particle phase space $\mathcal{S}$. Hence, by definition $\rho_c(\hat{\xi}^b):= \rho_c(s(\hat{\xi}^b))$. [Is this function $s$ the inverse of the coordinate mapping $\xi^b$?]. Then, since the physical state is independent of the coordinates, the state $s(\hat{\xi}^b)$ is the same state as $s(\xi^a) $ defined in terms of the original units $\xi^a$. Again, by definition, $\rho_c(s(\xi^a))=\rho_c(\xi^a) $. Since the distribution transforms as a density, it follows that $\rho_c(\xi^a)$ must equal the product of the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| $ times the density defined over the transformed units: $\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$. We can summarize the argument so far as a chain of equalities:

\begin{equation*}
\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)
\end{equation*}

\noindent
Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one. Assume for contradiction that there is only one state variable, so that there are no $k^i$ variables. Then we would have $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right| = 1$. But this would mean the unit change cannot be arbitrary. Therefore we must have two or more state variables, so that there must be at least one $k^i$ variable.

Next, we show that in fact there must be exactly one $k$ variable, conjugate to the single $q$ variable. Using the fact that the Jacobian determinant equals one, we determine how the state variables $k^i$ must change in order to compensate the change from $q$ to $\hat q$. Since $\hat{q}$ depends only on $q$, $\frac{\partial \hat{q}}{\partial k^a} = 0$. Hence, the Jacobian is an upper triangular block matrix:

\begin{equation}
%\begin{aligned}
|J| = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \begin{vmatrix}
\frac{\partial \hat{q}}{\partial q} & \frac{\partial \hat{q}}{\partial k^a} \\
\frac{\partial \hat{k}^b}{\partial q} & \frac{\partial \hat{k}^b}{\partial  k^a} 
\end{vmatrix} = \begin{vmatrix}
\frac{\partial \hat{q}}{\partial q} & 0 \\
\frac{\partial \hat{k}^b}{\partial q} & \frac{\partial \hat{k}^b}{\partial  k^a}
\end{vmatrix}
%\end{aligned}
\end{equation}

\noindent
The determinant of a triangular block matrix equals the product of determinants of the block matrices along the diagonal: $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| =  \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right|$. For the Jacobian determinant to equal one, we must therefore have $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| =   \left|\frac{\partial \hat{q}}{\partial q}\right|^{-1}$. Since $\hat q$ is a function only of $q$, $\frac{\partial \hat{q}}{\partial q}$ is a total derivative, so we obtain the inverse by swapping the numerator and denominator. Hence, we see that $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$. 
%  $\left|\frac{\partial \hat{q}}{\partial q}\right|^{-1} = \left|\frac{\partial q}{\partial \hat q} \right|$.

This equation provides only a single constraint, placed on the determinant of the transformation from $k^a$ to $\hat{k}^b$. Hence, if there were three or more $\xi^a$ variables, this constraint would be insufficient to recover the transformation of the $k^a$ uniquely. Yet in this case, $\hat q$  would not fully define the units for all other state variables, which is contrary to assumption. Consequently, there must be exactly two state variables: $q$ and a single $k$. So in this case of a single $q$ variable, coordinate-independent areas and densities can only be defined on a two-dimensional manifold.

\begin{comment}
\begin{equation}
\left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right| - \left|0\right| \left|\frac{\partial \hat{k}^b}{\partial q}\right|
= \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right|.
\end{equation}
 so $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$: 
\end{comment}

@@

We can now show that $\mathcal{S}$ must have at least the structure of a symplectic manifold. Invariant densities can only be assigned to infinitesimal areas. Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$, for $\Sigma \subset \mathcal{S}$. Because degrees of freedom are---at least locally---independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom. This means the volume form is given by $\omega^n$. Importantly, the volume form cannot be degenerate because otherwise there would be a volume in state space that could not contain states. But such a region of state space would---by definition---not be a volume. Therefore $\omega$ itself cannot be degenerate. As the degrees of freedom are independent, the number of states on a surface does not change if we translate it across independent degrees of freedom. If we imagine a parallelepiped, the integral over the surface must be zero (the integrals over opposite sides are equal and opposite) [explain/justify this more clearly. Is this coming from the two-form being antisymmetric?]. Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold. By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.

The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is a coordinate-invariant distribution over the state space $\mathcal{S}$ of its infinitesimal parts. If the state space of the particles is a manifold, then it must be a symplectic manifold: symplectic manifold structure is necessary for the distribution to be coordinate-invariant. 
% Which is the only type of manifold that supports coordinate-invariant distributions

\subsection{Deterministic and reversible evolution}
\label{deterministic}

@@

\begin{quotation}
\noindent
\textit{Deterministic and reversible evolution assumption}: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.

@@

[Should we footnote that this condition leaves out some classical systems, e.g. Norton's dome, Earman's space invaders?]

In our case, this applies to both the state of the whole system and to the state of its parts. Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle. Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$, at the initial time $t_0$, uniquely identifies the state of the system. Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial state at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e. the density remains the same throughout the evolution. In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa.  
%[could two different particles occupy the same initial state?].
%[what exactly does `initial particle' mean? Do we mean the initial state of the composite system or just a particular particle?]

Now consider the integral $\int_{\Sigma} \rho_c \omega(d\Sigma)$. The region $\Sigma$ will be mapped to a new region, but the fraction of the system found in both the old and new regions must be the same. Since both the integral and $\rho_c$ must remain constant during the evolution, the two-form $\omega$ must also remain the same. That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (otherwise, volumes would not be mapped to equal volumes, which would violate determinism). The evolution is therefore a symplectomorphism and corresponds to Hamiltonian evolution. Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution. In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.
